%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PMIx Standard Governance Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}

%%%%%%%%%%%%%%%%%%%
% Release Managers:
% - Set is_unofficial_draft = false to remove the watermark
% - Set VER to the correct version number
% - Set VERDATE to the Month+Year of the release
%%%%%%%%%%%%%%%%%%%
\usepackage[us,24hr]{datetime}
\usepackage{ifthen}
\newboolean{is_unofficial_draft}
\setboolean{is_unofficial_draft}{true}

% Text to appear in the footer on even-numbered pages:
\ifthenelse{\boolean{is_unofficial_draft}}
  {\newcommand{\VER}{1.6 (Draft)}
   \newcommand{\VERDATE}{\emph{Created on \today}}
  }
  {\newcommand{\VER}{1.6}
   \newcommand{\VERDATE}{MONTH DAY, YEAR}
  }
\newcommand{\footerText}{PMIx Standard Governance Rules -- Version \VER{} -- \VERDATE}

% Watermark for the Unofficial Drafts
\ifthenelse{\boolean{is_unofficial_draft}}
  {\usepackage{draftwatermark}
   \SetWatermarkText{\textbf{Unofficial Draft}}
   \SetWatermarkColor[gray]{0.85}
   \SetWatermarkFontSize{90pt}
   \SetWatermarkScale{1}
   \SetWatermarkAngle{45}
  }{}

%%%%%%%%%%%%%%%%%%%
% Links
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

%%%%%%%%%%%%%%%%%%%
% Verbatim code blocks
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\},fontsize=\small}
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\NormalTok}[1]{#1}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%%%%%%%%%%%%%%%%%%%
% Fonts
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{courier}
\usepackage{helvet}
\usepackage[utf8]{inputenc}
\usepackage{textgreek}

% Main body serif font:
\usepackage{tgtermes}
\usepackage[T1]{fontenc}

% Set default fonts:
\rmfamily\mdseries\upshape\normalsize

%%%%%%%%%%%%%%%%%%%
% Acronyms
\usepackage{acronym}
\acrodef{PMI}[PMI]{Process Management Interface}
\acrodef{PMIx}[PMIx]{Process Management Interface - Exascale}
\acrodef{ASC}[ASC]{Administrative Steering Committee}

%%%%%%%%%%%%%%%%%%%
% Footers
\usepackage{fancyhdr}     % makes right/left footers
\pagestyle{fancy}
\fancyhead{} % clear all header fields
\renewcommand{\headrulewidth}{0pt}

\cfoot{}
\lfoot{\bfseries \mdseries \footerText}
\rfoot{\bfseries \thepage}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}

\begin{titlepage}

\begin{center}
  \Huge \textsf{The PMIx Standard Governance Rules}

  \vspace{1.0in}

  \huge \textbf{Version \VER{}}

  \vspace{0.15in}

  \Large \textbf{\VERDATE}

\end{center}

\vspace{1.2in}

\vfill

\par
This document describes the PMIx Standard Governance Rules, version \VER{}.

\vspace{1em}

\textbf{Comments:}
Please provide comments on the PMIx Standard by filing issues on the document repository \url{https://github.com/pmix/governance/issues} or by sending them to the PMIx Community mailing list at \url{https://groups.google.com/forum/#!forum/pmix}.
Comments should include the version of this document that you are commenting about, and the page, section, and line numbers that you are referencing.
Please note that messages sent to the mailing list from an unsubscribed e-mail address will be ignored.

\vspace{1em}

Copyright\textsuperscript{\textcopyright} 2018-2021 PMIx \acf{ASC}.\\
Permission to copy without fee all or part of this material is granted,
provided the PMIx \ac{ASC} copyright notice and
the title of this document appear, and notice is given that copying is by
permission of PMIx \ac{ASC}.

\end{titlepage}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{pmix-standard}{%
\section{PMIx Standard}%
\label{pmix-standard}}

PMIx is an application programming interface (API) standard to provide
High Performance Computing (HPC) libraries and programming models with
portable and well defined access to commonly available distributed
computing system services.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{governance}{%
\subsection{Governance}%
\label{governance}}

The PMIx Standard is governed by an Administrative Steering Committee
(ASC) comprised of representatives from a wide range of research,
academic, and industrial organizations. Membership in the ASC is
determined by participation in the PMIx Standard meetings and a vote of
the ASC. Any organization meeting the membership criteria can become an
ASC \textit{Member}. An organization may refer to any entity, e.g., a
corporation, university, national laboratory, or an individual person
that is not employed by another Member organization (as detailed below).
Each Member shall be represented by a designated individual (the
\textit{Representative} who is authorized to vote on behalf of the Member.
If desired, a Member may designate an \textit{Alternate} who is authorized to
vote on behalf of the Member on occasions when the Representative is not
present.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{gov-asc-duties}{%
\subsubsection{ASC Duties and Organization}%
\label{gov-asc-duties}}

The ASC performs the following functions:

\begin{itemize}
\tightlist
\item
  Define the mission of the PMIx Standard.
\item
  Plan release timelines for versions of the PMIx Standard.
\item
  Vote on all issues requiring community consensus (e.g., acceptance of
  proposed modifications to the Standard), one vote per Member as cast
  by their Representative or Alternate. In addition to voting on
  proposed modifications, the ASC can be requested to vote on any issue
  submitted by a Member for consideration.
\item
  Assemble working groups to address specific areas of interest to the
  community.
\item
  Nominate and approve the appointment of individuals to fill the roles
  of Release Manager (per release) and Secretary, as described below.
\end{itemize}

The ASC is led by two \textit{Co-Chairs} responsible for organizing and
conducting the quarterly meetings. This includes:

\begin{itemize}
\tightlist
\item
  Announcing and finalizing meeting agendas.
\item
  Chairing the quarterly meetings
\item
  Ensuring timely publication of the minutes of the meeting, and
  tallying votes.
\end{itemize}

Co-Chairs are elected from amongst the ASC members during the last
quarterly meeting of each calendar year (using the voting system
described below). The newly elected Chairperson shall take office at the
beginning of the following calendar year, thus allowing overlap with the
outgoing Chairperson for efficient transition of active work items.
Co-Chairs serve for a period of two years, with seats filled on
alternate years - i.e., one seat will be elected during the last
quarterly meeting of each even year, and the other seat will be elected
during the last quarterly meeting of each odd year. There is no
restriction on the number of consecutive terms a given individual may
serve - however, it is expected and desirable that the positions rotate
amongst the Members. Likewise, there is no requirement that a Co-Chair
be the Representative or Alternate for their Member. Co-Chairs serving
as Representative or Alternate for a Member may vote on behalf of the
Member subject to the one vote per Member limit.

In the event a Co-Chair decides to step down from the position before
the completion of the two-year service period, the ASC will hold an
election for a replacement Co-Chair at the next quarterly meeting. The
newly elected replacement Co-Chair will serve for the remainder of the
time of the two-year service period of the original Co-Chair. The
replacement Co-Chair may seek election in subsequent terms.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{gov-asc-participation}{%
\subsubsection{ASC Participation}%
\label{gov-asc-participation}}

Participation in the consortium is both formal and informal:

\begin{itemize}
\item
  Informal participation is encouraged by working with the PMIx Standard
  community and/or one of the PMIx implementation libraries in an
  \emph{ad hoc} manner, such as (but not limited to) engaging in
  discussions on the mailing lists, reporting testing results, providing
  feature requests and bug reports, filing pull requests for proposed
  modifications, etc. \textit{Participants} do not have voting
  privileges on administrative matters, but are encouraged to attend the
  quarterly meetings and participate in those discussions. Organizations
  must attend meetings as Participants in order to qualify for elevation
  to full Member status, as described below.
\item
  Participants become eligible for elevation to Member status upon
  having at least one representative attend two quarterly meetings
  within the prior twelve months. All new Members must either request
  membership or be nominated by an existing ASC Member, and be approved
  by the ASC according to the voting rules described below. Membership
  confers representation on the ASC, thus granting the right to
  participate in formal votes, along with public acknowledgment for the
  entity as being a Member of the PMIx Standard ASC on the PMIx web
  site. Each Member organization shall appoint one official
  Representative with authority to vote on their behalf - if desired, an
  Alternate may also be designated. The Representative for an
  organization is responsible for notifying the ASC in advance of a
  temporary alternate should the Representative (and the Alternate, if
  designated) not be attending a meeting in order to confer voting
  rights to that individual.

  \begin{itemize}
  \tightlist
  \item
    Note: Individuals may also become Members so long as their direct
    employer is not a current Member of the ASC. For these purposes, a
    consultant is not considered a direct employee - thus, someone
    consulting for a current ASC Member is eligible to become an
    individual Member should they choose to do so. Individual Members
    who subsequently join an organization that is a Member will be
    absorbed into the larger organization's membership and will
    surrender their individual voting rights. Individuals requesting
    membership but are affiliated with multiple ASC Members can be
    granted individual Member status by vote of the ASC. This is
    intended to preclude undue influence in voting procedures by an
    organization with multiple individual Participants.
  \end{itemize}
\item
  A Member is removed from the ASC (a) upon having failed to have at
  least one representative attend two quarterly meetings in the prior
  twelve months, or (b) by vote of the other ASC members per the voting
  rules defined below. Attendance by a non-voting member of the
  organization (i.e., someone not officially designated as an alternate
  for voting purposes) counts towards meeting this requirement.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{gov-asc-meetings}{%
\subsubsection{ASC Meetings}%
\label{gov-asc-meetings}}

The ASC meets quarterly, with at least one quarterly meeting each year
to be held ``in-person'' at an agreed-upon location unless the ASC
Members vote otherwise. A vote to ``virtualize'' the in-person meeting
can only apply to that given year. Other quarterly meetings may be held
by teleconference in accordance with the desires of ASC Members.
Participants are welcome to attend any meeting in a non-voting capacity.
ASC quarterly meetings are conducted once per calendar quarter with a
preference for a time during the first month of the quarter. Precise
meeting dates are determined by the ASC Members in accordance with their
schedules, subject to the voting rules described below. Subsequent
changes to a given meeting's dates must be determined no later than two
quarterly meetings before the one being changed.

The Co-Chairs shall include links to each proposed modification to the
Standard up for consideration during that meeting in the published
agenda.

ASC quarterly meeting agenda items for reading and vote must be submitted to the Co-Chairs
and announced on the PMIx mailing list at least four weeks in advance of the meeting. Approximately four weeks in advance of
the meeting the Co-Chairs will announce the tentative agenda for review.
The community has until two weeks before the quarterly meeting to review and comment on
the agenda and PRs scheduled for vote or reading. During the review period, the contents of the agenda and PRs scheduled for vote or reading may be changed. The agenda will be finalized two weeks before the quarterly
meeting at which point no further additions or modifications will be accepted.
A PR posted for vote or reading in the ASC quarterly meeting must not have
any modification of the content of the branch after the agenda finalization
deadline. Comments on the PR are welcome at any time, but the content of the
PR must be frozen to allow for community review of the stable text. For a
formal reading or vote, it is required that a PDF rendering of the changed document
be sent to the PMIx mailing list and to the Co-chairs for distribution in the final agenda.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{modification-of-governance-rules}{%
\subsubsection{Modification of Governance Rules}%
\label{modification-of-governance-rules}}

Proposals to modify the PMIx governance document begin with the filing
of a GitHub Pull Request (PR) in a
\href{https://github.blog/2019-02-14-introducing-draft-pull-requests/}{Draft
PR} state and marked with a ``Governance'' label until ready for
submission for review. This PR shall serve as the vehicle for capturing
all discussion over the proposal. When ready for review, the author(s)
shall add the ``Proposed'' tag to the PR, thereby requesting that the
matter be placed on the agenda of the next quarterly meeting (subject to
the agenda finalization deadline). The \textit{Straw Poll} mechanism described
below is utilized to provide an initial screening of the proposal.
Passing of the Straw Poll \emph{does not} confer approval of the
proposal, but serves instead as a means for quickly pushing back
proposals that lack significant support. This is intended to both
protect the time of ASC Members and author(s), and to help guide
author(s) wishing to continue pursuit of the proposal.

Upon successfully completing the Straw Poll, the Co-Chairs shall label
the PR as ``Eligible'' to indicate that the proposal is
eligible for formal consideration by the ASC. Given the significance
attached to governance, formal approval of changes requires that the ASC
conduct a full presentation on the proposal at two consecutive quarterly
meetings. The timing of the second meeting may be delayed at the request
of the author(s) or ASC Members, subject to approval by the ASC. At the
conclusion of the first meeting, a formal vote of the ASC shall be held
(per the voting rules below) to determine if the proposal merits further
consideration. The Co-Chairs shall set a GitHub label indicating the
result as either ``First Vote Passed'' or ``Pushed
Back''.

At the conclusion of the second meeting where the proposal is up for
consideration, another formal vote of the ASC shall be held (per the
voting rules below) to determine final disposition of the proposal. The
Co-Chairs shall subsequently set a GitHub label indicating the result as
either ``Accepted'' or ``Pushed Back''. Accepted
proposals can be committed by the Co-Chairs in accordance with their
schedule, but take effect immediately (i.e., during the same meeting in
which they receive final approval).

Proposals that are pushed back at any point can be modified and
resubmitted by the author(s). However, the updated PR must pass through
the complete approval process from the beginning to ensure full
consideration of the revised request. A new PR for this purpose can be
submitted if the author(s) so desire, but must be linked to the original
PR in order to retain full information regarding prior concerns and
suggestions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{general-voting-rules}{%
\subsubsection{General Voting Rules}%
\label{general-voting-rules}}

The philosophy and goal of the ASC is to reach consensus on all, if not
most, issues. In the event that full consensus cannot be reached, the
ASC will use the following rules:

\begin{itemize}
\item
  Each active ASC Member will receive one vote. All votes are required
  to pass with 2/3 majority. A quorum (\textgreater50\%) of Members must
  be present (either physically or virtually). Abstentions will not be
  counted in the total votes when calculating whether there is a 2/3
  majority.
\item
  A Member may submit a vote \emph{in absentia} to the ASC Co-Chairs in
  advance of the meeting if attendance by their Representative or
  designated Alternate is not possible (e.g., a quarterly meeting that
  falls during a national holiday period), but each ASC Member can only
  cast one vote. This prevents a Member from voting on behalf of its own
  entity and acting as a proxy for another Member. \emph{In absentia}
  votes are counted by the Co-Chairs as if the Member were present (thus
  counting towards the meeting's quorum requirement), but are recorded
  as having been voted \emph{in absentia} in the official tally.
\end{itemize}

Revision Exception votes are special votes for PRs that have been modified
after the review period deadline before a quarterly meeting 
before a first vote or after having passed a first
vote. Revision Exception votes may be called by the proposer to evaluate 
whether the ASC is willing to proceed voting on the PR including the 
modifications. 
The idea behind Revision Exception votes is to maintain forward progress on
PRs even if changes are made after the quarterly meeting deadlines or after
a first vote.
Normally, changes acceptable for Revision Exception votes
are relatively minor and do not change the fundamentals of the PR. However,
if a quorum of ASC voting-eligible members all vote affirmatively  
for a Revision Exception vote, then the regular vote 
for the PR may proceed for the text of the PR including the modifications. 
A quorum of ASC voting-eligible members must be present and vote in the affirmative 
without any negative votes for the Revision Exception vote to pass.
If the Revision Exception vote does not pass, 
the proposer must choose to re-read the PR with the modifications
or to continue the normal vote without the modifications.

Straw Polls are used in several places as a means of gauging
community support for a proposal - e.g., in assessing acceptance for a
new Provisional API. Straw Polls are conducted on an informal basis
through the placement of emojis on a \textit{Straw Poll Comment} of an Issue or PR.
The poll comment consists of the following text:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Please use emoji reactions ON THIS COMMENT to indicate your position on this proposal.}

\BaseNTok{    * You do not need to vote on every proposal}
\BaseNTok{    * If you have no opinion, don't vote - that is also useful data}
\BaseNTok{    * If you've already commented on this issue, please still vote so}
\BaseNTok{          we know your current thoughts}
\BaseNTok{    * Not all proposals solve exactly the same problem, so we may end}
\BaseNTok{          up accepting proposals that appear to have some overlap}

\NormalTok{This is not a binding majority-rule vote, but it will be a very}
\NormalTok{significant input into the corresponding ASC decision.}

\NormalTok{Here are the meanings for the emojis:}

\BaseNTok{    * Hooray or Rocket: I support this so strongly that I}
\BaseNTok{          want to be an advocate for it}
\BaseNTok{    * Heart: I think this is an ideal solution}
\BaseNTok{    * Thumbs up: I'd be happy with this solution}
\BaseNTok{    * Confused: I'd rather we not do this, but I can tolerate it}
\BaseNTok{    * Thumbs down: I'd be actively unhappy, and may even consider}
\BaseNTok{          other technologies instead}

\NormalTok{If you want to explain in more detail, feel free to add another}
\NormalTok{comment, but please also vote on this comment.}
\end{Highlighting}
\end{Shaded}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{release-managers}{%
\subsubsection{Release Managers}%
\label{release-managers}}

One or more Release Managers are assigned per release series of the PMIx
Standard. Coordination and scheduling of releases in the series is
delegated to the Release Manager(s) by the ASC, subject to guidance
regarding needs on the part of Members. Release Managers have the
following duties:

\begin{itemize}
\tightlist
\item
  For each release, lead the establishment of release criteria by the
  ASC. This can include a list of proposed modifications to be decided
  upon prior to release and target release dates
\item
  Shepherd progress towards completing the schedule through interactions
  with proposal authors
\item
  Maintain the Revisions section of the document to ensure accurate
  communication of what has changed in each release
\end{itemize}

Release Managers shall be selected from within the community of Members,
but are not required to be a Representative or Alternate. Note that
Release Managers are not precluded from proposing modifications to the
Standard (including to the series under their management) during the
course of their assignment, nor are they prevented from voting if they
also serve as Representative or Alternate for their Member. In addition,
Release Managers are not responsible for promoting acceptance of any
proposal nor contributing to the completion of a proposal in order to
meet schedule, and may resolve any merge conflicts resulting from commit
of approved changes without further process.

The Release Managers shall notify the Co-Chairs when a minor release
candidate is complete and ready for publication. The Co-Chairs will add
an item to the agenda of the next quarterly meeting (subject to the
agenda finalization deadline) for final approval to publish the document. Once the ASC
has voted its approval, the Release Managers shall post the new release
on the community's web site and send an email announcing the release to
the community mailing list. Minor releases, therefore, can occur on an
at most a quarterly basis.

Major release candidates can be approved during any ASC quarterly
meeting. Release Managers must notify the Co-Chairs of a major release
candidate a minimum of two quarters before the target approval
meeting. The Co-Chairs will add an item to the agenda of each quarterly
meeting for discussion and review of the document, and on the target
meeting for vote on final approval. Once the ASC has voted its approval,
the Release Managers shall post the new release on the community's web
site and send an email announcing the release to the community mailing
list. Should the ASC not approve the release, subsequent release
candidates can be considered at upcoming ASC meetings. Note, no
subsequent major release may begin the approval process until after the
current candidate has been published. Major releases shall not occur
more than once per year.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{asc-secretaries}{%
\subsubsection{ASC Secretaries}%
\label{asc-secretaries}}

The ASC has two Secretary positions that are responsible for taking and publishing minutes of
quarterly meetings, and maintaining the PMIx Standard Web site. Other
duties may be defined by the ASC as required. Secretaries are elected
from amongst the ASC members during the last quarterly meeting of each
calendar year (using the voting system described above). The newly
elected Secretary shall take office at the beginning of the following
calendar year, thus allowing overlap with the outgoing Secretary for
efficient transition of active work items. ASC Secretaries serve for a
period of two years, with seats filled on alternate years - i.e., one
seat will be elected during the last quarterly meeting of each even
year, and the other seat will be elected during the last quarterly
meeting of each odd year. There is no restriction on the number of
consecutive terms a given individual may serve - however, it is expected
and desirable that the position rotate amongst the Members. Likewise,
there is no requirement that the Secretary be the Representative or
Alternate for their Member. A Secretary serving as Representative or
Alternate for a Member may vote on behalf of the Member subject to the
one vote per Member limit.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{asc-guidance}{%
\subsection{ASC Guidance}%
\label{asc-guidance}}

This section provides general guidance for the PMIx ASC and participants.
The rationale behind providing this guidance is to outline
philosophy and best practices the ASC should follow in 
maintaining the PMIx Standard. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{release-schedule-guidance}{%
\subsubsection{Release Schedule Guidance}%
\label{release-schedule-guidance}}

As described in Section~\ref{release-managers},
the schedule of major releases of the PMIx
Standard is determined by the Release Managers in coordination with other
members of the ASC. There is no prescribed major release schedule that the
Release Managers must follow. However, the Release Managers will prioritize
timely releases of substantive changes to the Stable entries in the 
PMIx Standard. Substantive changes
include: deprecated or removed
APIs, keys, or other text in the standard; alterations of APIs or keys; and
alterations of text that result in changed meaning or clarifications. The
Release Managers will prioritize major releases of the standard over waiting for
additional changes that are in progress but not yet accepted by the ASC.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{administrative-changes-guidance}{%
\subsubsection{Minor Document Changes Guidance}%
\label{administrative-changes-guidance}}

The intention of the rigorous process outlined in 
Section~\ref{the-standardization-process} for proposed
changes to the PMIx Standard is to ensure quality and consistency of 
the interface and the overall document. However, in some cases, 
PMIx participants may propose minor changes to the document that 
should not require the full process required for substantive changes. 
Examples of these minor changes include fixes for typographical errors 
(``typos''), missing or erroneous punctuation, and formatting changes.

In the case minor changes are proposed, the proposer should 
create an issue and pull request as described in Sections~\ref{initial-discussion}
and \ref{proposed-modification} and then bring the
changes to the attention of the ASC Co-Chairs and Release Managers.
The Co-Chairs and Release Managers will evaluate the proposed
changes to determine whether they are minor or substantive.
Additionally, all PMIx participants are encouranged to examine
and comment on the changes in the pull request to aid the
Co-Chairs and Release Managers in their decision.
If all Co-Chairs and Release Managers determine that the
changes are minor and do not alter the meaning of the current standard text, 
then the changes do not require the full process and
the changes can be applied to the document in the next release.
Alternatively, if any of the Co-Chairs or Release Managers determine 
that the changes do or could alter the meaning of the current standard text, 
the changes will be handled using the full rigorous process
as described in Section~\ref{the-standardization-process}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{the-standardization-process}{%
\subsection{The Standardization Process}%
\label{the-standardization-process}}

Contributions of any form (e.g., use cases, questions, suggestions) are
encouraged from anyone interested in the evolution of the PMIx Standard.
The community relies on Issues and Pull Requests to discuss changes to
the standard for a few reasons:

\begin{itemize}
\item
  It provides an archive of the conversation around a change to the
  standard that can be vital in understanding the rationale for the
  change if that conversation needs to be revisited in the future.
\item
  It facilitates asynchronous conversation from a geographically
  distributed user-base. To that end, any discussion on the
  teleconference about specific Issues or Pull Requests will be noted as
  a comment to that Issue or Pull Request.
\item
  The GitHub mechanism serves as a consensus building medium for
  accepting changes once all dissenting opinions and questions have been
  resolved, providing an archive of the responses to those concerns and
  those organizations participating in the process.
\end{itemize}

Templates are provided to assist in writing an Issue and/or Pull Request
that includes an appropriate level of information for that type of
request. Questions may be answered by anyone and participation from the
broad PMIx community is encouraged and welcomed, as are suggestions for
clarifications and corrections to the Standard. Feature requests are
best presented in the form of a use-case as this assists both the
community's understanding of the justification for providing the feature
and in assessing the feasibility/magnitude of effort required to provide
it.

Proposed modifications to the Standard can come in several forms:

\begin{itemize}
\item
  Addition of one or more new APIs and/or key-value attribute
  definitions
\item
  Changing the Standard classification of a current API and/or attribute
\item
  Deprecating an existing API and/or attribute
\item
  Clarifications and corrections to existing Standard language
\end{itemize}

The Standardization Process remains essentially the same regardless of
the type of modification being proposed, although the speed at which a
modification progresses through the process may vary considerably based
on the magnitude and controversial nature of the proposed change.
Discussions regarding a proposal are conducted either on the Issue/PR
itself or during the regular community teleconference (and captured as
notes on the Issue/PR). Teleconferences provide a high bandwidth
discussion format and valuable synchronization mechanism to guide
progress. Non-issue/PR related notes from the teleconferences are
archived on the PMIx community's
\href{https://github.com/pmix/pmix-standard/wiki\#meeting-information}{wiki}.

Note that a proposal can be withdrawn by the author(s) at any time for
any reason up to the time the proposal is committed to the Standard,
assuming acceptance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{classes-of-standard}{%
\subsubsection{Classes of Standard}%
\label{classes-of-standard}}

The PMIx Standard is composed of a combination of APIs and key-value
attributes. APIs are intentionally designed to be ``lightweight'' -
i.e., the API signature is generic with a minimal number of fixed
parameters in its signature. With few exceptions, APIs are required to
include an array of key-value attributes in their signature to
facilitate behavioral extensions without modifying the API signature.
Thus, PMIx attempts to minimize deprecation and modification of APIs as
use models evolve over time.

PMIx APIs and attributes are grouped into three classes:

\begin{itemize}
\item
  \textit{Provisional} entries in the Standard are officially part
  of the Standard but are not yet guaranteed to be stable - e.g., an API
  signature and/or attributes it supports may be changed, or the entry
  can be completely removed if subsequent evaluation proves negative.
  New provisional entries can be added to any minor release of the
  Standard. Subsequent changes to an entry requires warning in at least
  one minor release of the Standard document before acceptance - i.e., a
  proposed change to a provisional entry announced in version A.1 of the
  PMIx Standard may be committed in version A.2 of the Standard. This
  includes deprecation of an existing provisional entry. Thus, the
  intent is to provide a relatively easy mechanism by which proposals
  acceptable in principle to the community can enter the Standard prior
  to being fully stabilized, thereby allowing early adopters an
  opportunity to explore the proposed capabilities.

  All proposed API and attribute additions to the PMIx Standard enter the PMIx Standard
  under Provisional status. Attributes in the Provisional class may
  only be designated as ``optional'' by an API - i.e., an API definition
  cannot designate a Provisional attribute as ``required''. Users,
  however, are always free to ``require'' that an attribute be supported
  when calling the API, thus directing the API to respond with a ``not
  supported'' error should the attribute not be supported by it.
  Attributes that are intended to become designated as ``required'' by
  an API when elevated to Stable status shall be clearly designated
  as such in the proposal.

  Due to the flexible nature of the Provisional class, no guarantees are
  made with respect to backward or cross-version compatibility for
  entries in this class.
\item
  \textit{Stable} entries are permanent, non-changing parts of the
  PMIx Standard subject solely to deprecation. In this class, full
  guarantees are provided with respect to backward and cross-version
  compatibility beginning with the version upon which the entry was
  granted Stable status. No guarantees are made with respect to
  versions that included the entry during its Provisional status.

  Stable APIs may continue to define and/or adopt new supported
  key-value attribute pairs, thereby extending the \emph{behavior} of a
  stable API without modifying its signature. As previously stated,
  attributes in the Provisional class can only be specified as
  ``optional'' - therefore, \emph{required} attributes used by APIs in
  the Stable class must also reside in that class. A Stable API
  that wishes to utilize a Provisional attribute must either first
  seek elevation of the attribute to the Stable class so it can be
  ``required'', or must designate that attribute as ``optional''. This
  is necessary to ensure that all Stable parts of the Standard are
  capable of meeting the compatibility requirements. Provisional
  APIs are, of course, free to utilize any Stable attribute in
  addition to anything in the Provisional class.
\item
  \textit{Deprecated} entries are slated for removal in an upcoming
  major release of the Standard. Deprecation warning must be provided
  for a minimum of one major release - i.e., an entry that is marked for
  deprecation in release A.0 cannot be removed from the Standard until
  at least the A+2.0 release. Extensions to the deprecation deadline can
  be granted on a case-by-case basis upon vote of the ASC.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{initial-discussion}{%
\subsubsection{Initial Discussion}%
\label{initial-discussion}}

The process of amending the PMIx Standard begins with the opening of a
\textit{GitHub Issue} describing the proposed change. The Issue template
used and the degree of detail provided in the issue should be
commensurate with the magnitude of the proposed modification - e.g., a
spelling correction would use the ``Questions \& Clarification''
template and require very little justification, while a new API or
deprecation of an existing API might use the ``Use-Case'' template and
reasonably be expected to be accompanied by a more detailed
justification. Use of the template corresponding to the type of
modification being proposed is helpful to expedient processing of the
proposal - if in doubt as to the one to use, please ask.

The Issue provides a discussion forum for the contribution without
necessarily specifying precise changes to the PMIx Standard document.
Once an Issue is ready for discussion, it is helpful, though not
required, to send a message to the PMIx Forum mailing list to raise
awareness. Note that an Issue may not result in an actual change to the
Standard - e.g., it might serve simply as a focal point for community
discussion regarding a potential use-case or problem arising from a
use-case.

If a change to the PMIx Standard is included in the issue (or results
from the discussion in the issue), then a Pull Request must be created
(see below). Note that multiple Pull Requests may reference a single
Issue as authors consider alternative approaches to addressing the
Issue. The Issue will be closed once the associated PRs have been
resolved (either through by being accepted, withdrawn, or rejected) or
once the participants are satisfied with the conversation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{proposed-modification}{%
\subsubsection{Proposed Modification}%
\label{proposed-modification}}

When a specific change to the PMIx Standard has been identified a PR
must be opened with a reference made to the corresponding Issue(s). When
the PR is opened it should be placed in a
\href{https://github.blog/2019-02-14-introducing-draft-pull-requests/}{Draft
PR} state until the following conditions are met:

\begin{itemize}
\item
  The contents of the PR are ready for review and broad discussion with
  the PMIx community. Note that discussion over the PR is encouraged at
  all times even while in draft form.
\item
  An open-source implementation of the proposal is available and a
  reference to the implementation included in the PR. A complete
  implementation is not necessarily required to meet this condition -
  e.g., pseudocode detailing the execution path supporting the proposed
  modification can be provided. More complete prototype implementations
  may be requested by the PMIx community when proposals involving
  significant changes in behavior are made.
\end{itemize}

Only new commits are permitted on the branch once the
\href{https://github.blog/2019-02-14-introducing-draft-pull-requests/}{Draft
PR} designation has been removed. The branch associated with the PR
should not be squashed or force pushed after this point to preserve the
timeline of changes in relationship to the discussion that may occur on
the PR. Note that PRs that are squashed or force pushed remain eligible
for consideration, but may increase the difficulty of obtaining
consensus from the community.

Discussion of the PR will continue until all dissenting opinions and
questions have been addressed, upon which the PR may be put forward for
a formal decision by the ASC. Note that addressing dissenting opinions does
not necessarily mean achieving complete resolution on them. A dissent is
not considered to have ``veto'' power over a proposal, but must be
addressed to a degree that convinces the majority of ASC Members.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{provisional-acceptance}{%
\subsubsection{Provisional Acceptance}%
\label{provisional-acceptance}}

The process for acceptance to Provisional status begins when the
author(s) assign an ``RFC'' label to the GitHub PR, add the
Straw Poll Comment to it, and send an announcement of the
PR to the PMIx community mailing list requesting consideration of the PR
for Provisional status. The announcement shall include a link to the PR,
a brief description of the proposed change, and indicate any desired
timeline for approval. It is especially important that proposals tied to
contractual milestones be so designated to ensure that Members are
alerted to potential time pressures.

The Straw Poll Comment is used by the ASC as a mechanism for
obtaining a sense of how many people have viewed the PR and are
following the discussion. This serves no official purpose other than to
help define participation in the consensus building process and record
the tally of votes per emoji category, as defined above.

Anyone indicating concern must provide a comment on the ticket following
the Straw Poll Comment describing their concern. Sufficient
detail should be provided to allow the author(s) to respond to the
specific concerns. The author(s) can then work to address the concern,
modifying the PR as required - modification should be accompanied by an
appropriate comment citing the original participant who raised the
concern to ensure they are notified of the change. Participants and
Members shall indicate their acceptance or rejection of the modification
by so marking the change comment and adjusting their original ``vote''
on the Straw Poll if appropriate.

The ASC Co-Chairs shall monitor the status of discussion on the PR. Once
discussion appears to have consolidated (either measured by
participation in the Straw Poll, rate of comments, or a combination of
both), or at the request of the proposal's author(s), the Co-Chairs will
schedule the proposal for a reading at the next quarterly meeting of the
ASC. The proposal shall be included in the announced agenda for the
meeting along with a link to the PR and the results of the Straw Poll.

At least one proposal author must attend the quarterly meeting to
present the PR and participate in discussion regarding its approval. A
\textit{Reading} of the proposal shall be conducted at the meeting with the
author(s) reviewing any questions/concerns raised and how they were
addressed. During the presentation, someone designated by the Co-Chairs
will take notes on the PR (and/or Issue) on behalf of the presenter. 
After the reading in the same quarterly meeting, a formal vote of 
the ASC shall be held (per the above voting rules) to determine if
the proposal is acceptable as a provisional item. 
After the vote, the Co-Chairs shall set a GitHub Label corresponding to the
result - either indicating ``Accepted'' or ``Pushed Back'' based on the
results of the formal vote. Accepted proposals can be
committed to the next minor release version by the Release Manager in
accordance with their schedule. Proposals that are pushed back for
further work can be brought forward again at a later date, or can be
withdrawn by the author(s) by simply closing the PR.

\hypertarget{process-flow-for-provisional-proposals}{%
\paragraph{Process flow for Provisional
proposals.}\label{process-flow-for-provisional-proposals}}

In summary, the flow for proposing and gaining approval of a provisional
attribute and/or API consists of the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Open an Issue describing the proposed change to serve as a discussion
  forum
\item
  Open a Draft PR while writing the proposal itself
\item
  Remove the Draft PR status when ready for discussion
\item
  Discuss PR - only new commits can be added
\item
  Label the PR ``RFC'' and announce to the mailing list that it is ready
  for formal review
\item
  Add the Straw Poll Comment to request the PR be read at the
  next quarterly meeting
\item
  ASC Co-Chairs schedule proposal for Reading at next quarterly
  meeting
\item
  Author(s) present proposal for review at quarterly meeting
\item
  Formal ASC vote taken to determine if proposal is acceptable as  provisional 
\item
  ASC Co-Chairs mark proposal as either ``Accepted'' or ``Pushed Back''
\item
  Accepted proposals merged by Release Managers
\item
  Corresponding Issue closed
\end{enumerate}

Note that a provisional proposal could be accepted in as little as one quarterly
meeting.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{elevation-to-stable-status}{%
\subsubsection{Elevation to Stable Status}%
\label{elevation-to-stable-status}}

Provisional elements of the Standard that have been included in at least
two minor releases become eligible for elevation to Stable
status. The elevation process begins with an Issue requesting
consideration of the change, identifying the specific elements (APIs and
attributes) to be included, and a justification for Stable status.
Justification should be commensurate with the magnitude of the request -
e.g., elevating an API to Stable likely requires more justification
than a key-value attribute. While there is no clear formula for
justifying elevation, typical considerations might include the breadth
of adoption (by implementers, system management vendors, and/or
programming libraries) and stability of the elements themselves (i.e.,
time since last modification).

Note that attributes can be proposed for elevation separate from the
APIs that use them. Promotion of one or more attributes independent of
any API may, for example, occur when a Stable API has expressed a
desire to utilize them. In such cases, the proposal should clearly
identify the Stable APIs making the request and provide some
justification for it.

When ready for consideration, the author(s) shall create a corresponding
PR (linked to the Issue) in a
\href{https://github.blog/2019-02-14-introducing-draft-pull-requests/}{Draft
PR} state until ready for submission for review and label the PR with
the ``Elevate'' tag, thereby requesting that the matter be placed on the
agenda of the next quarterly meeting (subject to the agenda finalization
deadline). As with the process for Provisional acceptance, the
Straw Poll mechanism is utilized to provide an initial
screening of the proposal. Passing of the Straw Poll \emph{does not}
confer elevation to Stable status, but serves instead as a means for
quickly pushing back proposals that lack significant support. This is
intended to both protect the time of ASC Members and author(s), and to
help guide author(s) wishing to continue pursuit of the proposal.

Upon successfully completing the Straw Poll, the Co-Chairs shall label
the PR as ``Eligible'' to indicate that the proposal is
eligible for formal consideration by the ASC. Given the significant
commitment conferred by the PMIx community to Stable elements,
formal approval of elevation requires that the ASC conduct a full
presentation on the proposal at two consecutive quarterly meetings. The
timing of the second meeting may be delayed at the request of the
author(s) or ASC Members, subject to approval by the ASC. At the
conclusion of the first meeting, a formal vote of the ASC shall be held
(per the above voting rules) to determine if the proposal merits further
consideration. The Co-Chairs shall set a GitHub label indicating the
result as either ``First Vote Passed'' or ``Pushed
Back''.

At the conclusion of the second meeting where the proposal is up for
consideration, another formal vote of the ASC shall be held (per the
above voting rules) to determine final disposition of the proposal. The
Co-Chairs shall subsequently set a GitHub label indicating the result as
either ``Stable'' (indicating that the proposed elements have
been elevated to that class) or ``Pushed Back''. Accepted
proposals can be committed to the next major release version by the
Release Manager in accordance with their schedule.

Proposals that are pushed back at any point can be modified and
resubmitted by the author(s). However, the updated PR must pass through
the complete elevation process from the beginning to ensure full
consideration of the revised request. A new PR for this purpose can be
submitted for the identical elements but must be linked to the original
PR in order to retain full information regarding prior concerns and
suggestions.

\hypertarget{process-flow-for-elevation-proposals}{%
\paragraph{Process flow for Elevation
proposals.}\label{process-flow-for-elevation-proposals}}

In summary, the flow for proposing and gaining approval for elevation of
an attribute and/or API to Stable status consists of the following
steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Open an Issue describing the proposed change to serve as a discussion
  forum
\item
  Open a Draft PR while writing the proposal itself
\item
  Remove the Draft PR status when ready for discussion
\item
  Discuss PR - only new commits can be added
\item
  Label the PR ``Elevate'' and announce to the mailing list that it is
  ready for formal review
\item
  Add the Straw Poll Comment to request feedback
\item
  If it passes the Straw Poll, then ASC Co-Chairs label it as
  ``Eligible'' for consideration at next quarterly meeting
\item
  ASC Co-Chairs schedule proposal for Reading at next quarterly
  meeting
\item
  Author(s) present proposal for review at quarterly meeting
\item
  Formal ASC vote taken to designate as ``Pushed Back'' or ``First Vote
  Passed''
\item
  If passed, then author(s) present it again at a following quarterly
  meeting
\item
  Formal ASC vote taken to designate as ``Pushed Back'' or ``Accepted as
  Stable''
\item
  Accepted proposals merged by Release Managers
\item
  Corresponding Issue closed
\end{enumerate}

Note that a proposal requires a minimum of two quarterly meetings to be
accepted and that actual publication of the accepted proposal must be
included in the next major release of the standard.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{deprecation}{%
\subsubsection{Deprecation}%
\label{deprecation}}

Deprecation of one or more existing elements of the Standard, be they
APIs or attributes, starts by mirroring the path of a newly proposed
modification. An Issue proposing the deprecation and explaining the
reasons behind it shall be opened for initial discussion. Impact on
backward or cross-version compatibility must be identified in the
discussion to ensure adequate consideration is given to such issues.
When ready (either in parallel with the Issue or after an appropriate
amount of discussion on the Issue has concluded), the author(s) must
file a Deprecation PR containing the specific changes to the document
where deprecation warnings must be added to the Standard and/or wording
needs to be modified - examples are provided in the GitHub Deprecation
Template. The PR is opened in a
\href{https://github.blog/2019-02-14-introducing-draft-pull-requests/}{Draft
PR} state until ready for submission for review. When ready for
consideration, the author(s) shall label the PR with the ``Deprecate''
tag, thereby requesting that the matter be placed on the agenda of the
next quarterly meeting (subject to the agenda finalization deadline). The same
Straw Poll voting mechanism, including presentation at a
quarterly meeting, shall be used to determine if the proposal merits
deeper consideration by the ASC.

Deprecation of attributes can be proposed separately from the APIs that
support them - i.e., it is not required that an API which utilizes an
attribute be deprecated at the same time as the attribute. The process
of deprecating an attribute is, however, somewhat more complicated as it
may (a) be used in multiple Stable as well as Provisional APIs
and (b) impact the behavior of an API that supports it. Care must be
taken, therefore, to identify \emph{all} places in the Standard that are
impacted by the deprecation proposal. Where Stable APIs are
impacted, a deprecation warning for the change in \emph{behavior} must
be included along with warning of deprecation of the attribute itself.

Deprecation of Provisional elements proceeds along an accelerated
path. Upon initial acceptance based on the final results of the
Straw Poll, the Co-Chairs shall schedule a final formal
vote on the proposed deprecation at the next quarterly meeting. Thus, a
decision to deprecate an existing Provisional element requires both
a Straw Poll and a single formal vote by the ASC before being accepted.
The accepted deprecation proposal shall be labeled by the Co-Chairs as
``DEPRECATED'' and with the version of the Standard where the
deprecated elements shall be removed. Provisional elements must only
be marked as deprecated for a minimum of one \emph{minor} release of the
Standard before being removed - i.e., a provisional element that is
marked for deprecation in release A.1 can be removed from the Standard
in release A.2.

Deprecation of Stable elements follows a similar path, but requires
\emph{two} formal votes by the ASC before being accepted. This mirrors
the acceptance path for elevating the elements to Stable status. The
``double hurdle'' is intentional as removing an existing Stable
element of the Standard is considered to be a significant matter
potentially impacting a wide range of the PMIx community. The affected
elements shall remain in the Standard, marked for deprecation, for a
minimum of one \emph{major} release - i.e., an entry that is marked for
deprecation in release A.0 cannot be removed from the Standard until at
least the A+2.0 release.

Once officially deprecated, the PR carrying the deprecation warning(s)
shall be committed to the Standard by the Release Manager in accordance
with their schedule. Extensions to the deprecation cycle (i.e., the
length of time the elements must remain in the Standard marked as
``deprecated'') can be granted on a case-by-case basis upon vote of the
ASC.

At the conclusion of the deprecation cycle, a second ``removal'' PR
shall be submitted by the author(s) to actually remove the deprecated
element(s) from the PMIx Standard document. When ready for
consideration, the author(s) shall label the PR with the ``Removal''
tag, thereby requesting that the matter be placed on the agenda of the
next quarterly meeting (subject to the agenda finalization deadline). The Straw
Poll approval process utilized for Provisional proposals is
utilized to ensure adequate review is performed to avoid inadvertent
removal of elements not included in the original deprecation proposal.
Once the process has been concluded, the Release Manager shall commit
the change in accordance with their schedule.

\hypertarget{process-flow-for-deprecation-proposals}{%
\paragraph{Process flow for Deprecation
proposals.}\label{process-flow-for-deprecation-proposals}}

In summary, the flow for proposing and gaining approval for elevation of
an attribute and/or API to Stable status consists of the following
steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Open an Issue describing the proposed deprecation to serve as a
  discussion forum
\item
  Open a Draft PR while writing the proposal itself
\item
  Remove the Draft PR status when ready for discussion
\item
  Discuss PR - only new commits can be added
\item
  Label the PR ``Deprecate'' and announce to the mailing list that it is
  ready for formal review
\item
  Add the Straw Poll Comment to request feedback
\item
  If it passes the Straw Poll, then ASC Co-Chairs label it as
  ``Eligible'' for consideration at next quarterly meeting
\item
  ASC Co-Chairs schedule proposal for Reading at next quarterly
  meeting
\item
  Author(s) present proposal for review at quarterly meeting
\item
  Formal ASC vote taken

  \begin{itemize}
  \tightlist
  \item
    Provisional elements: designate as ``Pushed Back'' or
    ``Deprecated''
  \item
    Stable elements: designate as ``Pushed Back'' or ``First Vote
    Passed''

    \begin{itemize}
    \tightlist
    \item
      If passed, then author(s) present it again at a following
      quarterly meeting
    \item
      Formal ASC vote taken to designate as ``Pushed Back'' or
      ``Deprecated''
    \end{itemize}
  \end{itemize}
\item
  Accepted proposals merged by Release Managers
\item
  Corresponding Issue closed
\end{enumerate}

Note that a proposal to deprecate Provisional elements can be
approved in as little as one quarterly meeting, while a proposal to
deprecate Stable elements requires a minimum of two quarterly
meetings to be accepted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{major-text-changes}{%
\subsubsection{Major Text Changes}%
\label{major-text-changes}}

Major Text Changes are defined as significant changes to descriptive text in
the document such as introductory text in a chapter, definitions of terms,
re-organization of chapter contents or chapter placement, non-errata level
changes to descriptions of APIs including advice to implementors/users.
A major text change can target an upcoming minor or major release of the
standard, but significant changes in the semantics of existing
APIs shall target major releases. The definition of ``significant''
is determined by the members of the ASC.

The process for major text changes begins with a GitHub Draft PR while writing the proposal, optionally proceeded by a GitHub Issue for more general discussion and guidance.
When the GitHub PR is ready for discussion the author(s) must remove the Draft PR status, add the ``RFC'' and ``Major Text Change'' labels, and add the Straw Poll Comment to it.
Then the author(s) must send an announcement of the PR to the PMIx community mailing list requesting consideration of the PR as a major text change.
This starts the more formal discussion and Straw Poll voting process.

The Straw Poll Comment is used by the ASC as a mechanism for
obtaining a sense of how many people have viewed the PR and are
following the discussion. This serves no official purpose other than to
help define participation in the consensus building process and record
the tally of votes per emoji category, as defined above.
The author(s) must engage with any concerns raised on the PR in an attempt to
resolve those concerns before moving to the next stage in the process.

The ASC Co-Chairs shall monitor the status of discussion on the PR. Once
discussion appears to have consolidated (either measured by
participation in the Straw Poll, rate of comments, or a combination of
both), or at the request of the proposal's author(s), the Co-Chairs will
schedule the proposal for a reading at the next quarterly meeting of the
ASC. The proposal shall be included in the announced agenda for the
meeting along with a link to the PR. The PR shall explicitly specify if
it targets an upcoming minor or major release version of the standard.
The Co-Chairs shall label the PR as ``Eligible'' to indicate that the proposal is
eligible for formal consideration by the ASC at the next quarterly meeting.

Formal approval of a major text change requires that the ASC conduct a full
presentation on the proposal at two consecutive quarterly meetings. The
timing of the second meeting may be delayed at the request of the
author(s) or ASC Members, subject to approval by the ASC.

At least one proposal author must attend the quarterly meeting to
present the PR and participate in discussion regarding its approval. A
\textit{Reading} of the proposal shall be conducted at the meeting with the
author(s) reviewing any questions/concerns raised and how they were
addressed. During the presentation, someone designated by the Co-Chairs
will take notes on the PR (and/or Issue) on behalf of the presenter.
At the conclusion of the reading, a formal vote of the ASC shall be held
(per the above voting rules) to determine if the proposal merits further
consideration or pushed back for further modification.
The Co-Chairs shall set a GitHub label indicating the
result as either ``First Vote Passed' or ``Pushed Back''.

At the conclusion of the second meeting where the proposal is up for
consideration, another formal vote of the ASC shall be held (per the
above voting rules) to determine final disposition of the proposal. The
Co-Chairs shall subsequently set a GitHub label indicating the result as
either ``Accepted'' or ``Pushed Back''. Accepted
proposals can be committed to the next specified target release version
by the Release Manager in accordance with their schedule.

Proposals that are pushed back at any point can be modified and
resubmitted by the author(s). However, the updated PR must pass through
the complete major text change process from the beginning to ensure full
consideration of the revised request. A new PR for this purpose can be
submitted for the identical elements but must be linked to the original
PR in order to retain full information regarding prior concerns and
suggestions. Alternatively, the ASC may suggest a Revision Exception vote
instead of pushing back the PR. The ASC may also suggest a Revision Exception
vote to change the target release for the PR; in this case, the PR should
be rebased on the new target release with no other changes to its history.

\hypertarget{process-flow-for-major-text-changes}{%
\paragraph{Process flow for Major Text Changes}%
\label{process-flow-for-major-text-changes}}

In summary, the flow for proposing and gaining approval for a major text
change consists of the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Optionally, open an Issue describing the proposed change to serve as a discussion
  forum
\item
  Open a Draft PR while writing the proposal itself
\item
  When the PR is ready for discussion:
  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item 
    Remove the Draft PR status
  \item
    Label the PR ``RFC'' and ``Major Text Change''
  \item
    Milestone the PR with the target (minor or major) release of the standard
  \item
    Add the Straw Poll Comment to request feedback
  \item
    Send announcement to the mailing list that it is ready for formal review
  \end{enumerate}
\item
  Discuss PR - only new commits can be added
\item
  ASC Co-Chairs will schedule the proposal for Reading at next quarterly
  meeting and label the PR with ``Eligible''
\item
  Author(s) present proposal for review at quarterly meeting
\item
  Formal ASC vote taken to designate as ``Pushed Back'' or ``First Vote Passed''
\item
  If passed, then author(s) present it again at a following quarterly meeting
\item
  Formal ASC vote taken to designate as ``Pushed Back'' or ``Accepted``
\item
  Accepted proposals merged by Release Managers
\item
  Any corresponding Issue(s) are closed
\end{enumerate}

Note that a proposal requires a minimum of two quarterly meetings to be
accepted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{errata-changes}{%
\subsubsection{Errata Changes}%
\label{errata-changes}}

Errata changes are defined as ``small'' changes that correct errors or clarify ambiguities in the document.
The definition of ``small'' is determined by the members of the ASC.
As such, any errata item may be deemed by an ASC member as a major change (e.g., because it changes semantics in a subtle, but significant way) and thus need to go through a more formal process such as the \hyperlink{provisional-acceptance}{Provisional acceptance} process.
Note that for typos or grammar edits the \hyperlink{administrative-changes-guidance}{Minor Document Changes Guidance} applies.

The process for errata changes begins with a GitHub Draft PR while writing the proposal, optionally proceeded by a GitHub Issue for more general discussion and guidance.
When the GitHub PR is ready for discussion the authors(s) must remove the Draft PR status, add the ``RFC'' and ``Errata'' labels, and add the Straw Poll Comment to it.
Then the author(s) must send an announcement of the PR to the PMIx community mailing list requesting consideration of the PR as an errata change.
This starts the more formal discussion and Straw Poll voting process.

The Straw Poll Comment is used by the ASC as a mechanism for
obtaining a sense of how many people have viewed the PR and are
following the discussion. This serves no official purpose other than to
help define participation in the consensus building process and record
the tally of votes per emoji category, as defined above.
The author(s) must engage with any concerns raised on the PR in an attempt to
resolve those concerns before moving to the next stage in the process.

The ASC Co-Chairs shall monitor the status of discussion on the PR. Once
discussion appears to have consolidated (either measured by
participation in the Straw Poll, rate of comments, or a combination of
both), or at the request of the proposal's author(s), the Co-Chairs will
schedule the proposal for a reading at the next quarterly meeting of the
ASC. The proposal shall be included in the announced agenda for the
meeting along with a link to the PR.
The Co-Chairs shall label the PR as ``Eligible'' to indicate that the proposal is
eligible for formal consideration by the ASC at the next quarterly meeting.

At least one proposal author must attend the quarterly meeting to
present the PR and participate in discussion regarding its approval. A
\textit{Reading} of the proposal shall be conducted at the meeting with the
author(s) reviewing any questions/concerns raised and how they were
addressed. During the presentation, someone designated by the Co-Chairs
will take notes on the PR (and/or Issue) on behalf of the presenter.
At the conclusion of the reading, a formal vote of the ASC shall be held
(per the above voting rules) to determine if the proposal is accepted or
pushed back for further modification.
The Co-Chairs shall set a GitHub label indicating the
result as either ``Accepted' or ``Pushed Back''.

Accepted proposals can be committed to the next minor release version by
the Release Manager in accordance with their schedule.
Proposals that are pushed back for further work can be brought forward
again at a later date, or can be withdrawn by the author(s) by
closing the PR.

\hypertarget{process-flow-for-errata-changes}{%
\paragraph{Process Flow for Errata Changes.}%
\label{process-flow-for-errata-changes}}

In summary, the flow for proposing and gaining approval of a errata change consists of the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Optionally, open an Issue describing the proposed change to serve as a discussion
  forum
\item
  Open a Draft PR while writing the proposal itself
\item
  When the PR is ready for discussion:
  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item 
    Remove the Draft PR status
  \item
    Label the PR ``RFC'' and ``Errata''
  \item
    Add the Straw Poll Comment to request feedback
  \item
    Send announcement to the mailing list that it is ready for formal review
  \end{enumerate}
\item
  Discuss PR - only new commits can be added
\item
  ASC Co-Chairs will schedule the proposal for Reading at next quarterly
  meeting and label the PR with ``Eligible''
\item
  Author(s) present proposal for review at quarterly meeting
\item
  Formal ASC vote taken to designate as ``Pushed Back'' or ``Accepted''
\item
  Accepted proposals merged by Release Managers
\item
  Any corresponding Issue(s) are closed
\end{enumerate}

Note that an errata changes proposal could be accepted in as little as one quarterly
meeting.

\end{document}
